Information Set Generation in Kriegspiel

### 2010-12-14 paper/paper.tex
Reordered several elements and combined a table

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  Figure~\ref{abbrevoutput} shows the information set for black
after both player's have made five moves.  From this list of possibilities, black can infer that white must have a
bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with black's observations
up to that point. Tables~\ref{whitetimes} and ~\ref{blacktimes} show the size of the information sets for each player
for each move in the game and the amount of time to find the full set on a single processor.  Figure~\ref{verboseoutput}
shows the full output of the program for the same point in the game (five moves for each player) and illustrates all of
the possible positions in the belief state.

@@

\section{Appendix}
Figure~\ref{fulltree} shows the full game tree for a simple variant of poker known as Kuhn poker.  There are three
cards: a king, a queen, and a jack.  The dealer deals one card to each of two players.  Each player antes one unit, and
there is one simple round of betting, in which the size of the pot may be doubled.  Left branches denote check/fold;
right branches denote bet/call.  For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose
decision points are shown in triangles.  Connections between nodes in the same information set are shown with dotted
lines.

@@

Table~\ref{sampletimes}
\begin{table}
\begin{tabular}{ccc}
1 & 18 & 0.000\\
2 & 18 & 0.000\\
3 & 18 & 0.000\\
4 & 279 & 0.008\\
5 & 242 & 0.028\\
6 & 216 & 0.052\\
7 & 176 & 0.064\\
8 & 3406 & 0.092\\
9 & 2191 & 0.288\\
10 & 1423 & 0.508\\
11 & 1363 & 0.596\\
12 & 1416 & 0.744\\
13 & 1416 & 0.836\\
14 & 3440 & 0.97\\
15 & 3428 & 1.25\\
16 & 50521 & 1.82\\
17 & 43192 & 5.91\\
18 & 26128 & 7.94\\
19 & 19061 & 9.92\\
\end{tabular}
\caption{Solution counts and running times for a sample kriegspiel game}
\label{whitetimes}
\end{table}

Table~\ref{sampletimes}
\begin{table}
\begin{tabular}{ccc}
1 & 20 & .000\\
2 & 19 & .000\\
3 & 404 & .008\\
4 & 401 & .024\\
5 & 1472 & .044\\
6 & 155 & .096\\
7 & 293 & .116\\
8 & 158 & .128\\
9 & 118 & .148\\
10 & 25 & .152\\
11 & 798 & .164\\
12 & 518 & .192\\
13 & 13564 & .304\\
14 & 12394 & .776\\
15 & 343652 & 3.75\\
16 & 320704 & 17.3\\
17 & 490162 & 99.9\\
18 & 3792 & 119\\
19 & 14836 & 121 \\
\end{tabular}
\caption{Solution counts and running times for a sample kriegspiel game}
\label{blacktimes}
\end{table}

%\nocite{*}
\nocite{richards09information}
\nocite{russell05efficient}
\nocite{kuhn03lectures}
\nocite{kuhn97classics}
\nocite{parker05sampling}
\nocite{nance06reasoning}
\nocite{li94chess}

\bibliographystyle{acm}
\bibliography{paper}
%\begin{thebibliography}{99}
%\end{thebibliography}

%\begin{figure}
\begin{verbatim}
19
Moves:

**********************************************
R N B Q K B N R 
P P P P P P P P 
- - - - - - - - 
- - - - - - - - 
- - - - - - - - 
- - - - - - - - 
p p p p p p p p 
r n b q k b n r 
**********************************************
Failed moves: 0
from: 51 to: 35 blocked: 0 checked: 0

**********************************************
R N B Q K B N R 
P P P P P P P P 
- - - - - - - - 
- - - - - - - - 
- - - p - - - - 
- - - - - - - - 
p p p - p p p p 
r n b q k b n r 
**********************************************
Failed moves: 0
from: 8 to: 24 blocked: 0 checked: 0

**********************************************
R N B Q K B N R 
- P P P P P P P 
- - - - - - - - 
P - - - - - - - 
- - - p - - - - 
- - - - - - - - 
p p p - p p p p 
r n b q k b n r 
**********************************************
Failed moves: 0
from: 58 to: 30 blocked: 0 checked: 0

**********************************************
R N B Q K B N R 
- P P P P P P P 
- - - - - - - - 
P - - - - - b - 
- - - p - - - - 
- - - - - - - - 
p p p - p p p p 
r n - q k b n r 
**********************************************
Failed moves: 0
from: 9 to: 17 blocked: 0 checked: 0

**********************************************
R N B Q K B N R 
- - P P P P P P 
- P - - - - - - 
P - - - - - b - 
- - - p - - - - 
- - - - - - - - 
p p p - p p p p 
r n - q k b n r 
**********************************************
Failed moves: 1
from: 57 to: 42 blocked: 0 checked: 0

**********************************************
R N B Q K B N R 
- - P P P P P P 
- P - - - - - - 
P - - - - - b - 
- - - p - - - - 
- - n - - - - - 
p p p - p p p p 
r - - q k b n r 
**********************************************
Failed moves: 0
from: 10 to: 26 blocked: 0 checked: 0

**********************************************
R N B Q K B N R 
- - - P P P P P 
- P - - - - - - 
P - P - - - b - 
- - - p - - - - 
- - n - - - - - 
p p p - p p p p 
r - - q k b n r 
**********************************************
Failed moves: 1
from: 35 to: 27 blocked: 0 checked: 0

**********************************************
R N B Q K B N R 
- - - P P P P P 
- P - - - - - - 
P - P p - - b - 
- - - - - - - - 
- - n - - - - - 
p p p - p p p p 
r - - q k b n r 
**********************************************
Failed moves: 0
from: 1 to: 16 blocked: 0 checked: 0

**********************************************
R - B Q K B N R 
- - - P P P P P 
N P - - - - - - 
P - P p - - b - 
- - - - - - - - 
- - n - - - - - 
p p p - p p p p 
r - - q k b n r 
**********************************************
Failed moves: 0
from: 27 to: 19 blocked: 0 checked: 0

**********************************************
R - B Q K B N R 
- - - P P P P P 
N P - p - - - - 
P - P - - - b - 
- - - - - - - - 
- - n - - - - - 
p p p - p p p p 
r - - q k b n r 
**********************************************
Failed moves: 0
from: 13 to: 21 blocked: 0 checked: 0

**********************************************
R - B Q K B N R 
- - - P P - P P 
N P - p - P - - 
P - P - - - b - 
- - - - - - - - 
- - n - - - - - 
p p p - p p p p 
r - - q k b n r 
**********************************************
BEGINNING INFORMATION SET GENERATION
1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6

@@

\begin{figure}
%\begin{listing}
\begin{lstlisting}
void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
	uint16_t* possHistory, uint16_t** levels, int depth)
{
  if (!samePawnTries(trueState, possState, whiteMove)) return;  
  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
  if (depth == maxdepth) { // Then we have found a solution
	nSolutions++; return; // Report solution
  }
  uint16_t newPossState[16]; 
  int nMoves = 0;
  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
  assert (nMoves < NMOVES); //

Figure~\ref{codelisting}
\end{document}

>>>>

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  Figure~\ref{abbrevoutput} shows the information set for black
after both player's have made five moves.  From this list of possibilities, black can infer that white must have a
bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with black's observations
up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for each move in the game
and the amount of time to find the full set on a single processor.  Appendix B shows the full output of the program for
the same point in the game (five moves for each player) and illustrates all of the possible positions in the belief
state.

\begin{table}
\centering
\begin{tabular}{rrrrr}
 & \multicolumn{2}{c}{\bf White} & \multicolumn{2}{c}{\bf Black} \\
{\bf Ply} & {\bf Size} & {\bf Time} & {\bf Size} & {\bf Time} \\
1 & 18 & 0.000 & 20 & .000 \\
2 & 18 & 0.000 & 19 & .000 \\
3 & 18 & 0.000 & 404 & .008\\
4 & 279 & 0.008 & 401 & .024\\
5 & 242 & 0.028 & 1472 & .044\\
6 & 216 & 0.052 & 155 & .096\\
7 & 176 & 0.064 & 293 & .116\\
8 & 3406 & 0.092 & 158 & .128\\
9 & 2191 & 0.288 & 118 & .148\\
10 & 1423 & 0.508 & 25 & .152\\
11 & 1363 & 0.596 & 798 & .164\\
12 & 1416 & 0.744 & 518 & .192\\
13 & 1416 & 0.836 & 13564 & .304\\
14 & 3440 & 0.97 & 12394 & .776\\
15 & 3428 & 1.25 & 343652 & 3.75\\
16 & 50521 & 1.82 & 320704 & 17.3\\
17 & 43192 & 5.91 & 490162 & 99.9\\
18 & 26128 & 7.94 & 3792 & 119\\
19 & 19061 & 9.92 & 14836 & 121\\
\end{tabular}
\caption{Solution counts and running times for a sample kriegspiel game}
\label{bothtimes}
\end{table}

%Table~\ref{sampletimes}
%\begin{table}
%\begin{tabular}{ccc}
%1 & 20 & .000\\
%2 & 19 & .000\\
%3 & 404 & .008\\
%4 & 401 & .024\\
%5 & 1472 & .044\\
%6 & 155 & .096\\
%7 & 293 & .116\\
%8 & 158 & .128\\
%9 & 118 & .148\\
%10 & 25 & .152\\
%11 & 798 & .164\\
%12 & 518 & .192\\
%13 & 13564 & .304\\
%14 & 12394 & .776\\
%15 & 343652 & 3.75\\
%16 & 320704 & 17.3\\
%17 & 490162 & 99.9\\
%18 & 3792 & 119\\
%19 & 14836 & 121 \\
%\end{tabular}
%\caption{Solution counts and running times for a sample kriegspiel game}
%\label{blacktimes}
%\end{table}

\section{The Information Set Generation Algorithm}
\begin{figure}
%\begin{listing}
\begin{lstlisting}
void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
	uint16_t* possHistory, uint16_t** levels, int depth)
{
  if (!samePawnTries(trueState, possState, whiteMove)) return;  
  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
  if (depth == maxdepth) { // Then we have found a solution
	nSolutions++; return; // Report solution
  }
  uint16_t newPossState[16]; 
  int nMoves = 0;
  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
  assert (nMoves < NMOVES); //

Figure~\ref{codelisting}

@@

\section{Appendix A}
Figure~\ref{fulltree} shows the full game tree for a simple variant of poker known as Kuhn poker.  There are three
cards: a king, a queen, and a jack.  The dealer deals one card to each of two players.  Each player antes one unit, and
there is one simple round of betting, in which the size of the pot may be doubled.  Left branches denote check/fold;
right branches denote bet/call.  For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose
decision points are shown in triangles.  Connections between nodes in the same information set are shown with dotted
lines.

@@

%\nocite{*}
%\nocite{richards09information}
%\nocite{russell05efficient}
%\nocite{kuhn03lectures}
%\nocite{kuhn97classics}
%\nocite{parker05sampling}
%\nocite{nance06reasoning}
%\nocite{li94chess}

%\begin{thebibliography}{99}
%\end{thebibliography}

%\begin{figure}

\section{Appendix B}
\begin{verbatim}
1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6

@@

\bibliographystyle{acm}
\bibliography{paper}

\end{document}
